### IDSL.GOA

IDSL.GOA is based on the curated Gene Ontology Database. It performs GO set
enrichment analysis and works with InChIKey identifiers. The list can be loaded
into a GUI on the website (<https://goa.idsl.me/>). For furhter information see the Help Section" in the shiny app.

This code filters metabolite entries to retain unique "InChIKey" identifiers,
ensuring that duplicate metabolites with missing 'InChIKey ML' annotations are
excluded. Only significant features are selected. This pre-processing step is essential for reliable downstream
analyses, facilitating robust associations between unique metabolites and
clinical outcomes.

IDSL.GOA does not accept duplicate InChIKey entries. The first 14 characters of the InChIKey ID are reported, as IDSL.GOA uses only the two-dimensional structure.

```{r extract InChIKeys of regulated compounds for IDSL.GOA, echo=FALSE, results='markup',tidy=TRUE}
resolve_InChIKey <- function(df) {

  first14 <- function(x) substr(x, 1, 14)

  df <- df %>%
    mutate(
      InChIKey_ML  = safe_col(., "InChIKey ML"),
      InChIKey_db  = safe_col(., "InChIKey"),
      InChIKey_mzC = safe_col(., "InChIKey mzC"),

      ML14  = first14(InChIKey_ML),
      DB14  = first14(InChIKey_db),
      mzC14 = first14(InChIKey_mzC)
    )

  # global ML key list (14-char!)
  ml_keys <- df$ML14 %>%
    na.omit() %>%
    unique()

  df %>%
    mutate(

      is_ML = !is.na(ML14) & nzchar(ML14),

      fallback14 = coalesce(DB14, mzC14),

      # block fallback globally if ML connectivity exists
      fallback_blocked = fallback14 %in% ml_keys,

      InChIKey_GOA = case_when(
        is_ML            ~ ML14,
        fallback_blocked ~ NA_character_,
        TRUE             ~ fallback14
      )

    ) %>%
    arrange(InChIKey_GOA) %>%
    select(
      -InChIKey_ML, -InChIKey_db, -InChIKey_mzC,
      -ML14, -DB14, -mzC14,
      -fallback14, -fallback_blocked, -is_ML
    )
}

goa.df <- resolve_InChIKey(xlfile.df)

```

The merged InChIKey identifiers are stored in a text file for each comparison.

```{r write InChIKeys of regulated compounds for IDSL.GOA, echo=FALSE, results='markup',tidy=TRUE}

for (i in seq_along(log_cols)) {

  log_col  <- log_cols[i]
  pval_col <- pval_cols[i]

  print(log_col)
  print(pval_col)

  filtered_df <- goa.df %>%
    filter(
      (!!sym(log_col) >=  log2FC) |
      (!!sym(log_col) <= -log2FC) |
      ((!!sym(pval_col) <= pvalue) & (!!sym(log_col) >=  log2FCpV)) |
      ((!!sym(pval_col) <= pvalue) & (!!sym(log_col) <= -log2FCpV))
    ) %>%
    select(InChIKey_GOA, !!sym(log_col), !!sym(pval_col))

  # Remove actual NA and string "NA"
  inchi_vec <- filtered_df$InChIKey_GOA
  inchi_vec <- inchi_vec[!is.na(inchi_vec) & inchi_vec != "NA"]
 

  output_file_name <- format_output_filename(log_col)
  output_file_path <- file.path(
    GOA,
    paste0(user, "_GOA_", output_file_name, ".txt")
  )

  cat("###START###\n")
  cat("Nr. of rows IDSL.GOA:", output_file_name, length(unique(inchi_vec)), "\n")
  cat("###END###\n")

  write.table(
    unique(inchi_vec),
    file = output_file_path,
    sep = "\t",
    row.names = FALSE,
    col.names = FALSE,
    quote = FALSE
  )
}



```
