---
title: "CD-ID Selection: Export identifiers for different pathway and network analysis tools"
author:
  - "VBCF GmbH, Metabolomics"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 4
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    fig_caption: true
    df_print: paged
    keep_md: true
  pdf_document:
    toc: true
    toc_depth: '4'
    number_sections: true
    latex_engine: xelatex
params:
  path: "default"
  xlfile_df: NULL
  log2FC: 1
  log2FCpV: 0.75
  pvalue: 0.05
  user: "Project"
editor_options:
  markdown:
    wrap: 80
---

```{r knit setup, include=FALSE, message = FALSE, warning=FALSE, cache=FALSE}
# from advanced R - chapter debugging
# options(error = function() {
#   sink()
#   recover()
# })
#knitr::opts_chunk$set(cache=FALSE, 
#                      error = FALSE, 
                      # warning = FALSE, 
                      # message = FALSE, 
                      # cache.lazy = FALSE, 
                      # autodep = TRUE 
                      # )
# options(rlang_trace_top_env = rlang::current_env())
# options(error = function() {
#   sink()
#   print(rlang::trace_back(bottom = sys.frame(-1)), simplify = "none")
# })
```


```{r set working directory, echo=FALSE, results='markup', tidy=TRUE}
# Set a output directory
path <- params$path
#cat("The Zip file was saved to:", path)

```

# Data Preparation

The VBCF metabolomics core facility provides a result file that is suitable for the shiny app. Alternatively export a
file from Compound Discoverer 3.5 with the following columns (see "Help section" in shiny app):

-   Compound name, P-value, Log2 FC, Ratio, KEGG ID, HMDB ID, CSID, InChIKey

The file from VBCF contains only hits from the in-house database
and hits from mzCloud, a curated fragment library provided by Thermo Fischer
Scientific. If only data from our in-house library is preferred it can be sent
upon request or the excel file can be manually adjusted.

## Log2 fold changes and (adj.) p-value Input

```{r Define the values for log2 FC and p-values, echo=FALSE, results='markup', tidy=TRUE}
# adapt these values that it meets the requirements for your analysis

log2FC <- params$log2FC # log2 FC where no p-value is applied
cat("Log2FC no (adj.) p-value:", log2FC, "\n")

log2FCpV <- params$log2FCpV #log2 FC where p-value is applied
cat("Log2FC:", log2FCpV, "\n")

pvalue <- params$pvalue
cat("(adj.) p-value:", pvalue, "\n")

```

## Create and assign directories for output

A file structure is generated that assures that all output files are stored in
the correct file format for each pathway analysis tool.

```{r Define the directories and assign variables to each path, echo=FALSE, results='markup', tidy=TRUE}

subdir_names <- c("KEGG", "FELLA", "GOA", "Metscape", "MetaboAnalyst", "RAMPDB")

# Loop over each main subdirectory
for (subdir in subdir_names) {
  
  main_dir <- file.path(path, subdir)
  assign(subdir, main_dir)
  dir.create(main_dir, showWarnings = FALSE, recursive = TRUE)

  print(paste("Created directory and variable for:", subdir))
  #cat("Path:", subsub_dir, "\n\n")
}

# list variables
#ls()
# test if variable is a path
#cat("Path exists:", dir.exists(KEGG), "\n")
```

## Define the project name and import data

The excel file for pathway analysis is loaded and converted into a data frame.

```{r Adapting project name and import data, echo=FALSE, results='markup',tidy=TRUE}

# Change the user name to the project you want to run
user <- params$user
cat("Project name:", user, "\n")

# Import the Excel file
xlfile.df <- params$xlfile_df

if ("KEGG ID" %in% colnames(xlfile.df)) {
  xlfile.df[["KEGG ID"]] <- gsub(
    "^(cpd:|dr:|dr\\+|cpd\\+)",
    "",
    xlfile.df[["KEGG ID"]]
  )
}

head(xlfile.df)

```

## Store column names

Column names of "Log2 Fold Change", "P-value" or "Adj. P-value", and "Ratio" are
stored as global variables and will be further used to extract the relevant
information.

```{r Store global variables,echo=FALSE, results='markup', tidy=TRUE}

# Store Log2 Fold Change or Ratio and corresponding p-value columns
log_cols <- names(xlfile.df)[grepl("Log2 Fold Change:", names(xlfile.df))]

pval_cols <- sapply(log_cols, function(col) {
  p_col <- gsub("Log2 Fold Change:", "P-value:", col)
  adj_p_col <- gsub("Log2 Fold Change:", "Adj. P-value:", col)
  
  if (p_col %in% names(xlfile.df)) {
    p_col
  } else if (adj_p_col %in% names(xlfile.df)) {
    adj_p_col
  } else {
    NA
  }
})

pval_cols <- unname(pval_cols)


ratio_cols <- names(xlfile.df)[grepl("Ratio:", names(xlfile.df))]

pval_col_rat <- sapply(ratio_cols, function(col) {
  p_col <- gsub("Ratio:", "P-value:", col)
  adj_p_col <- gsub("Ratio:", "Adj. P-value:", col)
  
  if (p_col %in% names(xlfile.df)) {
    p_col
  } else if (adj_p_col %in% names(xlfile.df)) {
    adj_p_col
  } else {
    NA
  }
})

pval_col_rat <- unname(pval_col_rat)
```

```{r function for safe columns ,echo=FALSE, results='markup', tidy=TRUE}

safe_col <- function(df, col) {
  if (col %in% names(df)) df[[col]] else NA_character_
}

print(names(xlfile.df))
```

## File naming

Each file will be named according to the Log2 Fold Changes/Ratios that are
listed in the data frame. These are stored in the previous sections. The Prefix
as well as several special characters and white spaces will be removed or
replaced with this function.

```{r function for file naming, echo=FALSE, results='markup', tidy=TRUE}

format_output_filename <- function(log_col) {
  output_file_name <- gsub("Log2 Fold Change: ", "", log_col)  # Remove prefix
  output_file_name <- gsub("\\s+", "_", output_file_name)         # Replace whitespace with underscores
  output_file_name <- gsub("\\)", "", output_file_name)           # Remove ')'
  output_file_name <- gsub("\\(", "", output_file_name)           # Remove '('
  output_file_name <- gsub("/", "vs", output_file_name)           # Replace '/' with 'vs'
  output_file_name <- gsub(",", "", output_file_name)             # Remove ','
  return(output_file_name)
}
```

# Extract identifiers (IDs) for specific analysis tools

Each analysis tool requires a specific data format for the analysis. Therefore,
the data needs to be pre-processed and stored in the appropriate file format for
further handling.
