### MetaboAnalyst

In MetaboAnalyst pathway analysis, metabolite set
enrichment analysis (MSEA), and network analysis can be performed. The input data can have
various formats. HMDB and KEGG IDs can be mixed for ORA and will be used for the
analysis in MetaboAnalyst. For further information see the "Help Section" in the shiny app.

The list can be loaded into a GUI on the website
(<https://new.metaboanalyst.ca/ModuleView.xhtml>) or used in the R version. 
.
```{r extract data for MetaboAnalyst, echo=FALSE, results='markup',tidy=TRUE}

resolve_HMDBKEGG <- function(df) {

  df2 <- df %>%
    mutate(
      HMDB_ML   = safe_col(., "HMDB ML"),
      KEGG_ML   = safe_col(., "KEGG ML"),
      
      HMDB      = safe_col(., "HMDB ID"),
      KEGG_ID   = safe_col(., "KEGG ID"),
      
      InChI_ML  = safe_col(., "InChIKey ML"),
      InChI_db  = safe_col(., "InChIKey"),
      InChI_mzC = safe_col(., "InChIKey mzC"),
      InChI_use = coalesce(InChI_db, InChI_mzC)
    )
  
  blocked_mzC <- df2$InChI_use[!is.na(df2$InChI_ML)] %>%
                 na.omit() %>%
                 unique()

  used_ids <- character()

  hmdb_ml <- df2 %>%
    filter(!is.na(HMDB_ML)) %>%
    mutate(HMDBKEGG = paste0(HMDB_ML))

  used_ids <- hmdb_ml$HMDB_ML

  kegg_ml <- df2 %>%
    filter(is.na(HMDB_ML), !is.na(KEGG_ML), !(KEGG_ML %in% used_ids)) %>%
    mutate(HMDBKEGG = paste0(KEGG_ML))

  used_ids <- c(used_ids, kegg_ml$KEGG_ML)

  no_ml <- df2 %>%
    filter(
      is.na(HMDB_ML),
      is.na(KEGG_ML),
      !(InChI_use %in% blocked_mzC)
    )

  hmdb <- no_ml %>%
    filter(!is.na(HMDB), !(HMDB %in% used_ids)) %>%
    mutate(HMDBKEGG = paste0(HMDB))

  used_ids <- c(used_ids, hmdb$HMDB)

  kegg <- no_ml %>%
    filter(is.na(HMDB), !is.na(KEGG_ID), !(KEGG_ID %in% used_ids)) %>%
    mutate(HMDBKEGG = paste0(KEGG_ID))

  used_ids <- c(used_ids, kegg$KEGG_ID)

  bind_rows(hmdb_ml, kegg_ml,hmdb, kegg) %>%
    
    select(
      HMDBKEGG,
      everything(),
      -HMDB_ML, -KEGG_ML,
      -HMDB, -KEGG_ID,
      -InChI_ML, -InChI_use, -InChI_mzC, InChI_db
    ) %>%
    arrange(HMDBKEGG)
}

xlfile.MA.unique <- resolve_HMDBKEGG(xlfile.df)
head(xlfile.MA.unique)

```

The merged KEGG and HMDB identifiers are stored in a text file for each
comparison.

```{r write HMDB and KEGG IDs in txt file for MetaboAnalyst, echo=FALSE, results='markup',tidy=TRUE}

# Loop through each Log2 Fold Change column
for (i in seq_along(log_cols)) {
  log_col <- log_cols[i]        # Current Log2 Fold Change column
  pval_col <- pval_cols[i]        # Corresponding p-value column
  
  print(log_col)
  print(pval_col)
  
  # Filter the dataframe based on the conditions
  filtered_df <- xlfile.MA.unique %>%
    filter(
      (!!sym(log_col) >= log2FC ) |
      (!!sym(log_col) <= -log2FC) |
      ((!!sym(pval_col) <= pvalue) & (!!sym(log_col) >= log2FCpV)) |
      ((!!sym(pval_col) <= pvalue) & (!!sym(log_col) <= -log2FCpV))
      ) %>%
    select(HMDBKEGG, !!sym(log_col), !!sym(pval_col)) %>%
    filter(!str_detect(HMDBKEGG, "NA"))
  
  head(filtered_df)
  
  HMDBKEGG_unique <- unique(filtered_df$HMDBKEGG)

  output_file_name <- format_output_filename(log_col)
  output_file_name_txt <- paste0(output_file_name, ".txt") 
  
  # Construct the full path with the FELLA directory
  output_file_path <- file.path(MetaboAnalyst, paste0(user,"_MA_", output_file_name_txt)) 
  
  cat("\n###START###\n")
  cat("Nr. of rows MetaboAnalyst:", output_file_name, length(HMDBKEGG_unique))
  cat("\n###END###\n")
  
  # Write only the KEGG column to a text file in the FELLA directory
  write.table(HMDBKEGG_unique,
              file = output_file_path, sep = "\t",
              row.names = FALSE, 
              col.names = FALSE,
              quote = FALSE)
}
```
